---
title: "TITLE"
subtitle: "SUBTILE"
format: 
  docx:
    number-sections: true
    toc: true
  gfm:
    number-sections: true
  html:
    toc: true
execute:
  echo: false
  warning: false
bibliography: references.bib
csl: vancouver-brackets-no-et-al.csl
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: setup
library(epoxy)
library(tidyverse)
```

```{r}
#| label: load-data

## tbl-01-work-packages --------------------------------------------------

tbl01 <- read_csv(here::here("data/tables/tbl-01-work-packages.csv"))

wp_no <- tbl01$WP

wp_name <- tbl01$Name

## tbl-02-wp-activities-research-questions ------------------------------

tbl02 <- read_csv(here::here("data/tables/tbl-02-wp-activities-research-questions.csv"))

wp1_activities <- filter(tbl02, WP == "WP1") |> 
  pull(`Project Activity`) 

wp1_rqs <- filter(tbl02, WP == "WP1") |>
  pull(`Research Question`)

wp2_activities <- filter(tbl02, WP == "WP2") |>
  pull(`Project Activity`)

wp2_rqs <- filter(tbl02, WP == "WP2") |>
  pull(`Research Question`)

wp3_activities <- filter(tbl02, WP == "WP3") |>
  pull(`Project Activity`)

wp3_rqs <- filter(tbl02, WP == "WP3") |>
  pull(`Research Question`)

wp4_activities <- filter(tbl02, WP == "WP4") |>
  pull(`Project Activity`)

wp4_rqs <- filter(tbl02, WP == "WP4") |>
  pull(`Research Question`)

```

# Call Info

## Website

-   Measure 1: Calls for Field-Specific Actions
-   https://ethrat.ch/en/measure-1-calls-for-field-specific-actions/
-   Explore projects will allow early phase exploration, testing and prototyping of what an ORD practice could be like. They will enable starting to engage with a community without necessarily building up something definitive. The aim is to help our researchers to explore if ORD practices can be developed and how.

## Aims

Funding for "Explore projects" aims at:

1.  Specifying ORD standards.
2.  Prototyping ORD tools.
3.  Building Communities to engage in ORD practices.
4.  Any other means that help explore and develop ORD practices.

It is important that Explore projects are not self-centred but emphasize exchanges in the spirit of Open Science with other researchers, technical experts, and stakeholders.

## Application guidelines (ETH Domain)

-   https://ethrat.ch/wp-content/uploads/2023/10/Application-Guidelines_TrackA_Explore_2nd_call_FINAL.pdf
-   It is the declared goal to fund proposals that make tools (e.g., software; hardware; protocols; standards; benchmarks; documentation and training materials; networking and education events, activities and resources) available which enable, facilitate and foster sharing research data based on ORD principles. Preference will be given to projects which engage and are connected to a community. Therefore, applicants are strongly invited to demonstrate why their project represents an ORD practice that encourages good research in their field.

## Application guidelines (ETHZ & EPFL)

-   https://ethrat.ch/wp-content/uploads/2023/10/Application-guidelines-Explore-EPFL-ETHZ-applicants-2nd-call-LATEST.pdf

For applicants at ETHZ

-   In this Explore call, applicants at ETHZ must contribute 25% of the total budget from the resources of their lab/group/unit, either in-cash or in-kind. See the budgeting guidelines below for eligible contributions.

For applicants at EPFL

-   In this Explore call, applicants at EPFL can be funded for 100% of the total budget.

# Project Proposal

Your project proposal must not exceed 12 pages, excluding the bibliography (A4, Arial 11-point font, 1.5 line-spacing, 2-centimeter margins). Please address all points below.

# Abstract

**Instructions**

- maax 1500 characters
- needed on submission portal
- will be displayed publicly, see for example: https://ethrat.ch/wp-content/uploads/2022/11/ORD_Explore_call1_decisiontable_for-publication.pdf

# Keywords

**Instructions**

- 5 keywords that characterise your proposal (separated by commas)
- needed on submission portal

# Proposal full title

**Instructions**

Please include a project title (max. 12 words); this title will be used for all correspondence. Proposal Acronym (mandatory)

# Background and motivation

**Instructions**

Please explain which aim(s) of the ORD "contribute projects" you seek to address:

1.  Specifying ORD standards.
2.  Prototyping ORD tools.
3.  Building Communities to engage in ORD practices.
4.  Any other means that help explore and develop ORD practices.

It is important that Explore projects are not self-centred but emphasize exchanges in the spirit of Open Science with other researchers, technical experts, and stakeholders.


## openwashdata community

The openwashdata community has a home. We received funding for under the first call for Explore projects by the ETH Board and successfully build up a community of people that are interested in open data and open code in the greater Water, Sanitation, and Hygiene (WASH) sector.  

We have a build a website with 10 visitors daily, a monthly newsletter reaching X people, a community chat with 100 members, a 10 week data science programme with X graduates, and most importantly, a defined data publishing workflow for datasets and code. The goals of all work packages have been reached and in this next phase of the openwashdata project, we hope establish the foundations for long-term sustainability. 

## data science for openwashdata 

With an unexpected 222 sig-ups, we created a 10-week long data science program that significantly contribute to the build up of the community. Of the 2.5 hours allocated to each module, participants spent about 1/3 of the time in break-out rooms of two to three people, solving coding assignments together. This has lead to a cohort of people that is connected, and that has a common understanding of the data science workflow we implement for the community. We have observed that participants have started to work together on their homework assignments and exchange contact details to connect beyond the course. All graduates have published a complete data science report using GitHub Pages and shared underlying data in a GitHub repository.

We are eager to continue working with the participants of the course by offering trainings in the data publishing workflow and more advanced data science topics. Further, we will refine the content of the course to provide a second offering in the first quarter of a potentially funded proposal.

This course will continue to be a vehicle for growth of the community. We envisage that the course can eventually be taught by other members of the community.

## strategic partnerships

The first year of the project has provided us with an opportunity to observe potential partnerships with instituions and organisations. We believe that we could provide more direct support to organisations by placing a co-funded data steward directly within the organisation. This data steward would be funded 50% by a successful ORD proposal and 50% by the organisation. The data steward would be responsible for implementing the data publishing workflow within the organisation, and would be supported by the openwashdata community. Such partnerships will further help the Global Health Engineering group at ETH Zurich to establish itself as a leader in the field of open data and open code in the WASH sector, and together with partners bring ORD practices to the WASH sector.

We also believe that we need to establish a defined governance process and a sounding board for long-term success. The sounding board will be made up of members of the community and will provide feedback on the direction of the project. The governance process will define how decisions are made within the community. All of this will help to establish a long-term funding strategy for the openwashdata community.

## metadata standards

The first phase of the project successfully established a data publishing workflow that follows  FAIR data practices and adheres to the highest standards of rigor, transparency and reproducibility. All work is under git version control and public on Github. The includes the (unprocessed) submitted data, the code that is needed to get data from this stage to a of processed and analysis-read data, and the code that is needed to generate the figures and tables in the website that is generated to publish the data. The website is published using the R data package, which also enables detailed documentation and metadata for each dataset. An assigned DOI through Zenodo and the addition of all authors through an ORCID id makes the data citeable and provides a long-term archiving solution.

In this next phase, we want to explore how we can further increase the discoverability of the created data packages. We are looking at identifying a list of existing metadata standards and schemas that could be applicable for further publication and develop the necessary code to transform the data package into these standards. Currently, data can easily be used by the R community, and we also still provide CSV and XLSX formats, however a large proportion of the science community uses Python. We will explore how we can make the data packages more accessible to the Python community by protoyping a Python data packages for the pip and conda package managers. 

## Problem statement

The Global Health Engineering group has managed to establish a thriving ORD community in the WASH sector within the first year. ORD standards for data packages are specified, tools and platforms for ORD practices have been tested and selected, and a strong community has been built. 



The community has a home on the internet, a monthly newsletter, a community chat, and a 10-week data science programme. The community has a defined data publishing workflow and has published 10 datasets and 10 code repositories. The community has a defined governance process and a sounding board. The community has a long-term funding strategy. The community has a defined data

. The community has a home on the internet, a monthly newsletter, a community chat, and a 10-week data science programme. The community has a defined data publishing workflow and has published 10 datasets and 10 code repositories. The community has a defined governance process and a sounding board. The community has a long-term funding strategy. The community has a defined data


# ORD project plan

<!--# Edit work package names, activities and research questions in corresponding Google Sheet at: https://drive.google.com/drive/u/1/folders/1rcNgjQcII-V5N3Q5Z_KjM_6ghxFve20e -->

**Instructions**

Please describe your project design, including how it addresses the aim(s) of the ORD "Explore projects" described above. Please refer to the scientific literature as well as any preliminary, unpublished results that have informed your approach.

-   Which specific gap(s) in knowledge will your work address? Why is your approach innovative?
-   What questions will the project answer? What specific outcomes will you focus on?
-   Which strategy will you use to address the questions you seek to answer?

## `r paste0(wp_no[1], ": ", wp_name[1])`

### Goal

-   Describe the work concisely in 1 paragraph (80 to 120 words)

### Activities

```{epoxy}
{1:length(wp1_activities)}. {wp1_activities[1:length(wp1_activities)]}
```

### Research Questions

```{epoxy}
{1:length(wp1_rqs)}. {wp1_rqs[1:length(wp1_rqs)]}
```

### Aims addressed

-   Write a sentence on how the aims of the Explore call are addressed by this WP and it's corresponding activities

## `r paste0(wp_no[2], ": ", wp_name[2])`

### Goal

-   Describe the work concisely in 1 paragraph (80 to 120 words)

### Activities

```{epoxy}
{1:length(wp2_activities)}. {wp2_activities[1:length(wp2_activities)]}
```

### Research Questions

```{epoxy}
{1:length(wp2_rqs)}. {wp2_rqs[1:length(wp2_rqs)]}
```

### Aims addressed

-   Write a sentence on how the aims of the Explore call are addressed by this WP and it's corresponding activities

## `r paste0(wp_no[3], ": ", wp_name[3])`

### Goal

-   Describe the work concisely in 1 paragraph (80 to 120 words)

### Activities

```{epoxy}
{1:length(wp3_activities)}. {wp3_activities[1:length(wp3_activities)]}
```

### Research Questions

```{epoxy}
{1:length(wp3_rqs)}. {wp3_rqs[1:length(wp3_rqs)]}
```

### Aims addressed

-   Write a sentence on how the aims of the Explore call are addressed by this WP and it's corresponding activities

## `r paste0(wp_no[4], ": ", wp_name[4])`

### Goal

-   Describe the work concisely in 1 paragraph (80 to 120 words)

### Activities

```{epoxy}
{1:length(wp4_activities)}. {wp4_activities[1:length(wp4_activities)]}
```

### Research Questions

```{epoxy}
{1:length(wp4_rqs)}. {wp4_rqs[1:length(wp4_rqs)]}
```

### Aims addressed

-   Write a sentence on how the aims of the Explore call are addressed by this WP and it's corresponding activities

# Impact

**Instructions**

Please address these specific points:

-   How sustainable is the proposed project inside the ETH Domain?

-   To what extent will the planned project substantially advance the ORD field, or solve a critical outstanding problem in the targeted field(s)?

-   To what extent does the proposal support a collaborative approach, involving significant synergies, complementarities, or/and scientific added-value to achieve its objectives?

# Work Packages and milestones

**Instructions**

Please include a detailed Gantt chart or table outlining the timeline of your project, including key activities and major deliverables.

------------------------------------------------------------------------

The following Table X shows a basic gantt chart against the four work packages, including program activities. Column "Lead" abbreviations:

"Any publications derived from this program will be published as open access material, following ORD practices and Open Science standards for computational reproducibility and sharing of data and code under FAIR principles."

Table X: https://docs.google.com/spreadsheets/d/18-6BHa8NR-g2Qlpu9yVSOEFXmsh6mRsAGjvWZMhBN4I/edit#gid=0

# Resources (including project costs)

Table Y: https://docs.google.com/spreadsheets/d/1MJkJN739eI9apa0Wfg1cHmIQgSshE1arCHBPWeFJwDk/edit#gid=0

# Budget justification and description of co-financing

Table Z: https://docs.google.com/spreadsheets/d/1Z7qNNp1DVB6F38eKfvx8FEUXzNi-ZPVfjtAetBJufdI/edit#gid=0

# Bibliography
