---
title: "TITLE"
subtitle: "SUBTILE"
format: 
  docx:
    number-sections: true
    toc: true
  gfm:
    number-sections: true
  html:
    toc: true
execute:
  echo: false
  warning: false
bibliography: references.bib
csl: vancouver-brackets-no-et-al.csl
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: setup
library(epoxy)
library(tidyverse)
```

```{r}
#| label: load-data

## tbl-01-work-packages --------------------------------------------------

tbl01 <- read_csv(here::here("data/tables/tbl-01-work-packages.csv"))

wp_no <- tbl01$WP

wp_name <- tbl01$Name

## tbl-02-wp-activities-research-questions ------------------------------

tbl02 <- read_csv(here::here("data/tables/tbl-02-wp-activities-research-questions.csv"))

wp1_activities <- filter(tbl02, WP == "WP1") |> 
  pull(`Project Activity`) 

wp1_rqs <- filter(tbl02, WP == "WP1") |>
  pull(`Research Question`)

wp2_activities <- filter(tbl02, WP == "WP2") |>
  pull(`Project Activity`)

wp2_rqs <- filter(tbl02, WP == "WP2") |>
  pull(`Research Question`)

wp3_activities <- filter(tbl02, WP == "WP3") |>
  pull(`Project Activity`)

wp3_rqs <- filter(tbl02, WP == "WP3") |>
  pull(`Research Question`)

wp4_activities <- filter(tbl02, WP == "WP4") |>
  pull(`Project Activity`)

wp4_rqs <- filter(tbl02, WP == "WP4") |>
  pull(`Research Question`)

```

# Call Info

## Website

-   Measure 1: Calls for Field-Specific Actions
-   https://ethrat.ch/en/measure-1-calls-for-field-specific-actions/
-   Explore projects will allow early phase exploration, testing and prototyping of what an ORD practice could be like. They will enable starting to engage with a community without necessarily building up something definitive. The aim is to help our researchers to explore if ORD practices can be developed and how.

## Aims

Funding for "Explore projects" aims at:

1.  Specifying ORD standards.
2.  Prototyping ORD tools.
3.  Building Communities to engage in ORD practices.
4.  Any other means that help explore and develop ORD practices.

It is important that Explore projects are not self-centred but emphasize exchanges in the spirit of Open Science with other researchers, technical experts, and stakeholders.

## Application guidelines (ETH Domain)

-   https://ethrat.ch/wp-content/uploads/2023/10/Application-Guidelines_TrackA_Explore_2nd_call_FINAL.pdf
-   It is the declared goal to fund proposals that make tools (e.g., software; hardware; protocols; standards; benchmarks; documentation and training materials; networking and education events, activities and resources) available which enable, facilitate and foster sharing research data based on ORD principles. Preference will be given to projects which engage and are connected to a community. Therefore, applicants are strongly invited to demonstrate why their project represents an ORD practice that encourages good research in their field.

## Application guidelines (ETHZ & EPFL)

-   https://ethrat.ch/wp-content/uploads/2023/10/Application-guidelines-Explore-EPFL-ETHZ-applicants-2nd-call-LATEST.pdf

For applicants at ETHZ

-   In this Explore call, applicants at ETHZ must contribute 25% of the total budget from the resources of their lab/group/unit, either in-cash or in-kind. See the budgeting guidelines below for eligible contributions.

For applicants at EPFL

-   In this Explore call, applicants at EPFL can be funded for 100% of the total budget.

# Project Proposal

Your project proposal must not exceed 12 pages, excluding the bibliography (A4, Arial 11-point font, 1.5 line-spacing, 2-centimeter margins). Please address all points below.

# Abstract

**Instructions**

- maax 1500 characters
- needed on submission portal
- will be displayed publicly, see for example: https://ethrat.ch/wp-content/uploads/2022/11/ORD_Explore_call1_decisiontable_for-publication.pdf

# Keywords

**Instructions**

- 5 keywords that characterise your proposal (separated by commas)
- needed on submission portal

# Proposal full title

**Instructions**

Please include a project title (max. 12 words); this title will be used for all correspondence. Proposal Acronym (mandatory)

# Background and motivation

**Instructions**

Please explain which aim(s) of the ORD "contribute projects" you seek to address:

1.  Specifying ORD standards.
2.  Prototyping ORD tools.
3.  Building Communities to engage in ORD practices.
4.  Any other means that help explore and develop ORD practices.

It is important that Explore projects are not self-centred but emphasize exchanges in the spirit of Open Science with other researchers, technical experts, and stakeholders.


## openwashdata community

The openwashdata community has a home. We received funding for under the first call for Explore projects by the ETH Board and successfully build up a community of people that are interested in open data and open code in the greater Water, Sanitation, and Hygiene (WASH) sector.  

We have a build a website with 10 visitors daily, a monthly newsletter reaching X people, a community chat with 100 members, a 10 week data science programme with X graduates, and most importantly, a defined data publishing workflow for datasets and code. The goals of all work packages have been reached and in this next phase of the openwashdata project, we hope establish the foundations for long-term sustainability. 

## data science for openwashdata 

With an unexpected 222 sig-ups, we created a 10-week long data science program that significantly contribute to the build up of the community. Of the 2.5 hours allocated to each module, participants spent about 1/3 of the time in break-out rooms of two to three people, solving coding assignments together. This has lead to a cohort of people that is connected, and that has a common understanding of the data science workflow we implement for the community. We have observed that participants have started to work together on their homework assignments and exchange contact details to connect beyond the course. All graduates have published a complete data science report using GitHub Pages and shared underlying data in a GitHub repository.

We are eager to continue working with the participants of the course by offering trainings in the data publishing workflow and more advanced data science topics. Further, we will refine the content of the course to provide a second offering in the first quarter of a potentially funded proposal.

This course will continue to be a vehicle for growth of the community. We envisage that the course can eventually be taught by other members of the community.

## strategic partnerships

The first year of the project has provided us with an opportunity to observe potential partnerships with instituions and organisations. We believe that we could provide more direct support to organisations by placing a data steward directly within these organisation. The data steward will be responsible to work with the organisation on identifying current ORD practices, develop a draft data management strategy, identify datasets that can readily be published openly, and identify datasets that need more careful consideration for publishing because of previous ethics approvals, data privacy, or industry partnerships. The openwashdata team will directly support this data steward with guidance and technical support in all activities. 

Such partnerships will further help the Global Health Engineering group at ETH Zurich to establish itself as a leader in the field of open data and open code in the WASH sector, and together with partners bring ORD practices to the WASH sector.

We also believe that we need to establish a defined governance process and a sounding board for long-term success. The sounding board will be made up of members of the community and will provide feedback on the direction of the project. The governance process will define how decisions are made within the community. All of this will help to establish a long-term funding strategy for the openwashdata community.


### UKZN WASH R&D Center

The UKZN WASH R&D Center is a research center at the University of KwaZulu-Natal in Durban, South Africa, established in 1971 and formerly known as the Pollution Research Group. The center is a leader in WASH R&D and has provided a platform for testing of innovative sanitation solutions for local and international service providers. The so-called Sanitation Prototype Field-Testing Platform is designed to expose sanitation technologies to real world conditions early in their development by testing prototypes in informal settlements and rural households in Durban, South Africa. Between 2017 and 2020, nearly 20 prototype technologies have been tested on this platform. It was a multi-discplinary partnership between university partners, local consultancy firm Khanyisa Projects, and with involvement of the local eThekwini Municipality's Water and Sanitation Unit (EWS). On a national level this work was supported by the Water Research Commission (WRC) and financially supported by the Bill and Melinda Gates Foundation (BMGF). 

Based on the experiences of prototyping various sanitation technologies, SASTEP (the South African Sanitation Technology Enterprise Programme) was created. The programme was created by the Water Research Commission (WRC) inpartnership with the Department of Science and Innovation (DSI), and the Bill and Melinda Foundation (BMGF) and with the support of the Department of Water and Sanitation (DWS). The WASH R&D Center is the research partner for the implementation of the programme. 

The WASH R&D team consists of more than 30 people, including two directorate staff, five administrative staff, four associate professors, four senior researchers, six post-doc researchers, five PhD students, and about ten MSc students. The research methods includes a combination of qualitative and quantitative research community research, fundamental laboratory research, and a field and test lab for industry partners. 

TODO: Figure out what that group of international laboratories is.

Over the years, the WASH R&D Center has collected a large amount of data. This includes data from the Sanitation Prototype Field-Testing Platform, but also data from the research community. The data is currently stored in a variety of locations, including the university's network drives, the university's data repository, and on the personal computers of researchers. The data is not published openly and is not easily accessible to the research community.

### BASEflow Malawi

BASEflow is a malawian social enterprise (non-profit organisation) that was founded in 2017 with the aim of improving water security in Malawi. The organisation is based in Blantyre, Malawi and has a team of X people. They have shown leadership in data collection and data sharing in the WASH sector in Malawi and support research institutions and government agencies with data collection and data management. During the first phase of openwashdata, they have published a dataset using the openwashdata workflow and are interested in continuing to work with us and develop a data management strategy for their organisation that enabled all data that can shared to be published openly.

This unique approach is rare in the WASH sector as non-academic organisations have less incentives to share their data. However, the data holds great value and is often sought after by government and researchers. BASEflow can set a precedent for other organisations in the sector to follow this approach and be recognised for it as a leader in the sector.

## metadata standards

The first phase of the project successfully established a data publishing workflow that follows FAIR data practices and adheres to the highest standards of rigor, transparency and reproducibility. All work is under git version control, public on Github and archived on Zneodo. This includes the (unprocessed) submitted data, the code that is needed to get data from this stage to a stage of processed and analysis-read data, and the code that is needed to generate the figures and tables in articles on websites that are created to publicize and curate the data. The website is published using the R data package ecosystem, which provides boilerplate templates for detailed documentation and provision of metadata for each dataset. An assigned DOI through Zenodo and the addition of all authors through an ORCID id makes the data findable and reusable.

In this next phase, we want to explore how we can further increase the findability of the created data packages. We are looking at identifying a list of existing metadata standards and schemas that could be applicable for further publication and we develop the necessary code to transform the data package into these standards (e.g. JSON Table Schema of Frictionless Standards, Dublin Core Metadata Element Set).

Currently, data can easily be used by the R community, and we also still provide CSV and XLSX formats, however a large proportion of the science community uses Python programming. We will explore how we can make the datasets more interoperable and accessible to the Python community by protoyping a Python data packages for the pip and conda package managers. 

Lastly, the R data packages are currently installed from Github using the `remotes` or `devtools` R packages. The effort of submitting all datasets to the Comprehensive R Archive Network (CRAN) is too time intensive to be justified at this point. However, we will explore to develop an R package and submit it to CRAN for providing a way to explore all published datasets and install them through this package. A publication of an R package on CRAN will significantly increase the findability of our data and further improve the process of exploring and reusing the data.

## Problem statement

The Global Health Engineering group has managed to establish a thriving ORD community in the WASH sector within its first year. ORD standards for data packages are specified, tools and platforms for ORD practices have been tested and selected, and a strong and active community has been built. 

In this next phase, we are hoping to set the foundation for long-term sustainability of the community and further increase the application of FAIR principles to the data that is published. This next phase will enable us to explore and develop partnerships, which the Global Health Engineering group will further leverage beyond the project's duration. 


# ORD project plan

<!--# Edit work package names, activities and research questions in corresponding Google Sheet at: https://drive.google.com/drive/u/1/folders/1rcNgjQcII-V5N3Q5Z_KjM_6ghxFve20e -->

**Instructions**

Please describe your project design, including how it addresses the aim(s) of the ORD "Explore projects" described above. Please refer to the scientific literature as well as any preliminary, unpublished results that have informed your approach.

-   Which specific gap(s) in knowledge will your work address? Why is your approach innovative?
-   What questions will the project answer? What specific outcomes will you focus on?
-   Which strategy will you use to address the questions you seek to answer?

## `r paste0(wp_no[1], ": ", wp_name[1])`

### Goal

-   Describe the work concisely in 1 paragraph (80 to 120 words)

### Activities

```{epoxy}
{1:length(wp1_activities)}. {wp1_activities[1:length(wp1_activities)]}
```

### Research Questions

```{epoxy}
{1:length(wp1_rqs)}. {wp1_rqs[1:length(wp1_rqs)]}
```

### Aims addressed

-   Write a sentence on how the aims of the Explore call are addressed by this WP and it's corresponding activities

## `r paste0(wp_no[2], ": ", wp_name[2])`

### Goal

-   Describe the work concisely in 1 paragraph (80 to 120 words)

### Activities

```{epoxy}
{1:length(wp2_activities)}. {wp2_activities[1:length(wp2_activities)]}
```

### Research Questions

```{epoxy}
{1:length(wp2_rqs)}. {wp2_rqs[1:length(wp2_rqs)]}
```

### Aims addressed

-   Write a sentence on how the aims of the Explore call are addressed by this WP and it's corresponding activities

## `r paste0(wp_no[3], ": ", wp_name[3])`

### Goal

-   Describe the work concisely in 1 paragraph (80 to 120 words)

### Activities

```{epoxy}
{1:length(wp3_activities)}. {wp3_activities[1:length(wp3_activities)]}
```

### Research Questions

```{epoxy}
{1:length(wp3_rqs)}. {wp3_rqs[1:length(wp3_rqs)]}
```

### Aims addressed

-   Write a sentence on how the aims of the Explore call are addressed by this WP and it's corresponding activities

## `r paste0(wp_no[4], ": ", wp_name[4])`

### Goal

-   Describe the work concisely in 1 paragraph (80 to 120 words)

### Activities

```{epoxy}
{1:length(wp4_activities)}. {wp4_activities[1:length(wp4_activities)]}
```

### Research Questions

```{epoxy}
{1:length(wp4_rqs)}. {wp4_rqs[1:length(wp4_rqs)]}
```

### Aims addressed

-   Write a sentence on how the aims of the Explore call are addressed by this WP and it's corresponding activities

# Impact

**Instructions**

Please address these specific points:

-   How sustainable is the proposed project inside the ETH Domain?

-   To what extent will the planned project substantially advance the ORD field, or solve a critical outstanding problem in the targeted field(s)?

-   To what extent does the proposal support a collaborative approach, involving significant synergies, complementarities, or/and scientific added-value to achieve its objectives?

# Work Packages and milestones

**Instructions**

Please include a detailed Gantt chart or table outlining the timeline of your project, including key activities and major deliverables.

------------------------------------------------------------------------

The following Table X shows a basic gantt chart against the four work packages, including program activities. Column "Lead" abbreviations:

"Any publications derived from this program will be published as open access material, following ORD practices and Open Science standards for computational reproducibility and sharing of data and code under FAIR principles."

Table X: https://docs.google.com/spreadsheets/d/18-6BHa8NR-g2Qlpu9yVSOEFXmsh6mRsAGjvWZMhBN4I/edit#gid=0

# Resources (including project costs)

Table Y: https://docs.google.com/spreadsheets/d/1MJkJN739eI9apa0Wfg1cHmIQgSshE1arCHBPWeFJwDk/edit#gid=0

# Budget justification and description of co-financing

Table Z: https://docs.google.com/spreadsheets/d/1Z7qNNp1DVB6F38eKfvx8FEUXzNi-ZPVfjtAetBJufdI/edit#gid=0

# Bibliography
